Challenges in Developing Great Quasi-Monte Carlo Software

### 2023-03-09 MCQMC2022ChallengesQMCSoftwareArticle.tex
SC: A few more minor changes

A QMC approximation of $\mu := \EE[f(\bsX)]$, $\bsX \sim \calU[0,1]^d$ can be implemented in a few steps:
\begin{enumerate}
    \item Draw a sequence of $n$ low-discrepancy (LD) \cite{Nie92, SloJoe94, DicEtal14a, DicEtal22a} nodes,
    $\bsx_1,\dots,\bsx_n$ that mimic $\calU[0,1]^d$.
    \item Evaluate the integrand $f$ at these nodes to obtain $f(\bsx_i)$, $i=1,\dots,n$.
    \item Estimate the true mean, $\mu$, by the sample mean,
    \begin{equation}
        \hat{\mu} := \frac{1}{n} \sum_{i=1}^n f(\bsx_i).
        \label{eq:mc_average}
    \end{equation}
\end{enumerate}
However, the practice of QMC is often more complicated.  The original problem may need to be rewritten to fit the above form and/or to facilitate a good approximation with small $n$.  Practitioners may wish to determine $n$ adaptively to satisfy a prescribed error tolerance.

@@

Recent interest in quality scientific software is exemplified by the 2021 US Department of Energy \emph{Workshop on the Science of Scientific-Software Development and Use} \cite{ASCR-SSSDU,osti_1846008}. This workshop not only discussed diagnostics and treatments for the challenges of developing great scientific software, but also emphasized the importance of implementing theoretical advancements into well written, accessible software libraries. Three cross-cutting themes arose in this workshop:
\begin{itemize}
    \item  We need to consider both human and technical elements to better understand how to improve the development and use of scientific software.

@@

LD generators are the most prevalent generally available components of QMC software.  These generators appear in Association of Computing Machinery  publications \cite{BraFox88,BraFoxNie92,HonHic00a}, FinDer \cite{PasTra95,FinDer}, libseq \cite{FriKel02,FriKelweb}, BRODA \cite{BRODA20a}, NAG \cite{NAG27}, MATLAB \cite{MAT9.13}, SamplePack \cite{SamplePack}, Gr\"unschlo{\ss}'s website \cite{GruWeb}, the Magic Point Shop \cite{NuyWeb}, R \cite{QRNG2020}, Julia \cite{QMCJulia}, PyTorch \cite{paszke2019pytorch}, SAS \cite{SAS_LD}, SciPy \cite{virtanen2020scipy}, and TensorFlow \cite{tfqfQMC2021a}, MatBuilder \cite{paulin2022}, and  QMCPy \cite{QMCPy2020a}.

@@

\subsection{Integrands and Variable Transformations} \label{CDHJS:sec:integrands}
The original integral or expectation arising from the application is defined in terms of the integrand, $g$, and the non-negative weight, $\lambda$.  For example,
\begin{itemize}
	\item $g$ is the discounted payoff of a financial derivative and $\lambda$ is the probability density function (PDF) for a discretized Brownian motion \cite{Gla03},

@@

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{cantileveredbeamwords.eps}
    \includegraphics[width=0.7\textwidth]{ldparallelbeam.eps}
    \caption{The expected deflection of a cantilevered beam with random Young's modulus.  QMCPy computes the answer in serial and in parallel using UM-Bridge and docker containers.  For small error tolerances, the time and sample size is $\calO(n^{-1})$ \SCComment{typo:$\calO(\varepsilon^{-1})$}, and parallel processing gives a speed-up of factor four to five. These plots may be reproduced using the Jupyter notebook \SCComment{(1) Add url. (2) What is the meaning of min and sec on the y-axis of lower left graph (The label says (s), which means seconds right?)? Are there numerical values on (left and right) y-axis? (3) The font size of the axis labels for the lower graphs could be made larger more like the top graph. (4) I guess the blue colors are covered by the orange colors in the bottom right graph. Maybe we can use blue crosses and orange broken lines instead.}}
    \label{fig:beam}
\end{figure}

>>>>

A QMC approximation of $\mu := \EE[f(\bsX)]$, $\bsX \sim \calU[0,1]^d$ can be implemented in a few steps:
\begin{enumerate}
    \item Draw a sequence of $n$ low-discrepancy (LD) \cite{Nie92, SloJoe94, DicEtal14a, DicEtal22a} nodes,
    $\bsx_1,\dots,\bsx_n$ that mimic $\calU[0,1]^d$.
    \item Evaluate the integrand $f$ at these nodes to obtain $f(\bsx_i)$, $i=1,\dots,n$.
    \item Estimate the true mean, $\mu$, by the sample mean,
    \begin{equation}
        \hat{\mu} := \frac{1}{n} \sum_{i=1}^n f(\bsx_i).
        \label{eq:mc_average}
    \end{equation}
\end{enumerate}
However, the practice of QMC is often more complicated.  The original problem may need to be rewritten to fit the above form and/or to facilitate a good approximation with small $n$.  Practitioners may wish to determine $n$ adaptively to satisfy a prescribed error tolerance. \SCComment{Use $\hat{\mu}_n$ in (1) so that it is consistent with (3).}

@@

Recent interest in quality scientific software is exemplified by the 2021 US Department of Energy \SCComment{report,} \emph{Workshop on the Science of Scientific-Software Development and Use} \cite{ASCR-SSSDU, osti_1846008}. This workshop not only discussed diagnostics and treatments for the challenges of developing great scientific software, but also emphasized the importance of implementing theoretical advancements into well written, accessible software libraries. Three cross-cutting themes arose in this workshop:
\begin{itemize}
    \item  We need to consider both human and technical elements to better understand how to improve the development and use of scientific software.

@@

LD generators are the most prevalent generally available \SCComment{do we need `generally available'?} components of QMC software.  These generators appear in Association of Computing Machinery  publications \cite{BraFox88,BraFoxNie92,HonHic00a}, FinDer \cite{PasTra95,FinDer}, libseq \cite{FriKel02,FriKelweb}, BRODA \cite{BRODA20a}, NAG \cite{NAG27}, MATLAB \cite{MAT9.13}, SamplePack \cite{SamplePack}, Gr\"unschlo{\ss}'s website \cite{GruWeb}, the Magic Point Shop \cite{NuyWeb}, R \cite{QRNG2020}, Julia \cite{QMCJulia}, PyTorch \cite{paszke2019pytorch}, SAS \cite{SAS_LD}, SciPy \cite{virtanen2020scipy}, and TensorFlow \cite{tfqfQMC2021a}, MatBuilder \cite{paulin2022}, and  QMCPy \cite{QMCPy2020a}.

@@

\subsection{Integrands and Variable Transformations} \label{CDHJS:sec:integrands}
The original integral or expectation arising from the \SCComment{an?} application is defined in terms of the integrand, $g$, and the non-negative weight, $\lambda$.  For example,
\begin{itemize}
	\item $g$ is the discounted payoff of a financial derivative and $\lambda$ is the probability density function (PDF) for a discretized Brownian motion \cite{Gla03},

@@

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{cantileveredbeamwords.eps}
    \includegraphics[width=0.7\textwidth]{ldparallelbeam.eps}
    \caption{The expected deflection of a cantilevered beam with random Young's modulus.  QMCPy computes the answer in serial and in parallel using UM-Bridge and docker containers.  For small error tolerances, the time and sample size is $\calO(n^{-1})$ \SCComment{typo:$\calO(\varepsilon^{-1})$}, and parallel processing gives a speed-up factor of four to five. These plots may be reproduced using the Jupyter notebook \SCComment{(1) Add url. (2) What is the meaning of min and sec on the y-axis of lower left graph (The label says (s), which means seconds right?)? Are there numerical values on (left and right) y-axis? (3) The font size of the axis labels for the lower graphs could be made larger more like the top graph. (4) I guess the blue colors are covered by the orange colors in the bottom right graph. Maybe we can use blue crosses and orange broken lines instead.}}
    \label{fig:beam}
\end{figure}
